Are historical stage records useful to decrease the uncertainty of flood frequency analysis ? A 200-year long case study

### 2022-12-14 ArticleChProp/main.tex
14-12 version, Jérôme suggestions

\usepackage{dsfont}
\usepackage{multirow}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{subcaption}

@@

\title{Are historical stage records useful to decrease the uncertainty of flood frequency analysis ? A 200-year long case study}
\author{Mathieu LUCAS, Benjamin RENARD, Jérôme LE COZ, Michel LANG, A. BARD, G. PIERREFEU}

@@

    This paper presents a chain of uncertainty estimation methods for flood frequency analysis in the context of long discharge series. Uncertainties are estimated and propagated from stage and rating curves to extreme flood quantiles estimates using Monte Carlo procedures. This method allows exploring the role of both streamflow and sampling uncertainties in design floods estimations and provides insights on how historical data may or may not improve design floods estimates. The procedure is applied to the 205-year long continuous discharge series of the Rhône River at Beaucaire (95 590 m\textsuperscript{2}). Total uncertainty on flood quantile is significantly reduced when the length of the series increases from 50 to 100 years. Beyond this length, the total uncertainty remains stable, despite 19th century flood discharges have larger uncertainties than “modern’ flood discharges (20th century and present). Addition of the two largest known floods in 1840 and 1856 remains critical, as the estimate of Q100 and Q1000 floods is revised upwards.

@@

    Rating curve uncertainty is also a major issue when dealing with streamflow series. Transforming stage into discharge requires calibration data (gaugings) to establish the stage/discharge relationship. Gaugings uncertainty depends on the measurement method (\citet{lecoz_quantification_2014}b and \citet{puechberty_charte_2017}). Moreover, the rating curve is also affected by uncertainties coming from the imperfection of the chosen model to represent the actual hydraulic configuration and from parameter estimation. Many methods have been proposed to quantify these uncertainties (\citet{petersen-overleir_bayesian_2009}; \citet{juston_rating_2014}; \citet{le_coz_combining_2014}; \citet{morlot_dynamic_2014}; \citet{coxon_novel_2015}; \citet{mcmillan_rating_2015}; \citet{mansanarez_rapid_2019}). A comparison of several of these methods has been recently proposed by \citet{kiang_comparison_2018}. Another important issue affecting streamflow data accuracy is rating changes. The stage/discharge relationship is frequently affected by changes coming from various origins that can be natural or anthropic, for instance: bed geometry evolution during floods or river works, aquatic vegetation growth and decay, ice cover... A regular monitoring through gaugings is essential to detect those changes \citep{ibbitt_gauging_1987}, that can be transient or sudden. Several methods have been proposed to deal with rating changes: estimating rating curves on moving temporal windows (\citet{westerberg_stage-discharge_2011} and \citet{guerrero_temporal_2012}), computing as many rating curves as there are gaugings \citep{morlot_dynamic_2014}, exploring changes in the annual minimum stages \citep{lapuszek_methods_2015}, selecting 0.5-years return period discharge as a threshold for rating changes \citep{mcmillan_impacts_2010}. More recently, \citet{darienzo_detection_2021} proposed a method based on a recursive segmentation procedure, accounting for both gaugings and rating curve uncertainties. This method has a particular interest when dealing with old and thus uncertain gaugings. Following the detection of rating shifts, rating curves should be estimated for each stability period. This task may not be straightforward, as the number of gaugings available within a stability period is not always sufficient to properly estimate the stage/discharge relationship for the whole discharge range. A common way to address this problem is to artificially repeat some gaugings from other stability periods (\citet{mcmillan_benchmarking_2012}; \citet{puechberty_charte_2017}). \citet{mansanarez_shift_2019} proposed an alternative approach to deal with this issue. They developed a stage-period-discharge (SPD) model  where some parameters of the rating curve are allowed to vary across periods, while the others are supposed fixed. This method has the advantage of transferring information between periods to improve the rating curve estimation for all the stable stage/discharge periods, even when few gaugings are available.

@@

    \paragraph{}
    Once the stability periods have been identified, the next step is to estimate the rating curve associated with each period. \citet{mansanarez_shift_2019} developed a stage-period-discharge (SPD) model "based on the physical interpretation of changes in the stage-discharge relation across a series of stability periods". The SPD model is based on the BaRatin model \citep{le_coz_combining_2014}.
    \paragraph{} BaRatin uses the Bayesian paradigm to estimate the parameters of the general rating curve equation, a combination of power equations: $Q = a(h-b)^c$, where $Q$ is the discharge in m\textsuperscript{3}/s, the flow depth is the difference between the stage $h$ (in m) and an offset $b$, and $a$ and $c$ being the coefficient and the exponent of the power function. The rating curve equation is deduced from a hydraulic analysis of the gauging station, aimed at identifying the main hydraulic controls governing the stage-discharge relation. The multiple controls can be activated successively or simultaneously. Bayesian inference allows deriving the posterior distribution of rating curve parameters by combining hydraulic information (priors for parameters of each hydraulic controls) and information from gaugings with uncertainty (likelihood).
    \paragraph{}
    The SPD model is estimating the rating curves of each stability based on the same principle, and is considering that several identified parameters are varying in time, throughout the stability periods. An important step is the identification of those varying parameters based on hydraulic analysis of the site. Generally, offsets or channel widths are suspected to change. The so-called local changes are supposed to affect the lowest control only (for instance the movement of the controlling riffle), whereas the global changes affect various controls at the same time (for instance, the scouring or filling of the channel, affecting low flows and main channel offsets at the same time). One way to follow the evolution of riverbed elevation and to help specifying priors for rating curve vertical evolution is to study the long-term evolution of the yearly lowest stages, as described by \citet{lapuszek_methods_2015}. See \citet{mansanarez_shift_2019} for a detailed description of prior specification for time-varying rating curves.

@@

    \paragraph{}    
    Thirteen shift times are detected at Beaucaire Restitution. As can be seen in figure \ref{fig:SegmBoth} (right), gauging frequency is far higher than for Pont de Beaucaire station (except during the 5 first years), allowing a better determination of the rating shift times. 
    Due to the lack of gaugings at the beginning of the series, only one rating shift is detected but many shifts potentially took place in those first four years as morphological readjustments and dredging occurred (see previous section). This first shift is assumed to be assigned to the first important flood of the station in 1976, after which the channel is stabilized. The following shift occurred during 1994 flood, one of the biggest of the station. The most notable flood at Beaucaire Restitution occurred in 2003 (11 500 m\textsuperscript{3}/s, with a return period of about 100 years according to \citet{medd_debit_2005}). Unsurprisingly, the stage/discharge relationship is considerably disrupted by this event, as reflected by the six rating shifts detected from 2003 to 2005. Two out of these six shifts were discarded because the shift amplitude is considered minor based on further analysis of the corresponding rating curve change. The largest flood within posterior pdf of those shifts is almost always corresponding to the 2003 flood. This is also the case for 2005, 2008 and 2009 shifts, for which the posterior pdf spans many years and include 2003 flood. Therefore, the shift dates are assumed to be located to the maxpost shift times, as several shifts can not be located at the same date. Lastly, the last shift of 2019 is also discarded because the shift amplitude is considered minor based on further analysis. Finally, ten rating shifts were retained. This brings the number of stability periods to eleven for Beaucaire Restitution, as seen in table \ref{tab:ShiftDates}.

@@

    \paragraph{}
    Uncertain rating curves are estimated using \citet{mansanarez_shift_2019} SPD model, for each stability period detected previously. For Pont de Beaucaire, this leads to ten rating curves that show a good adequacy with gaugings (figure \ref{subfig:RcPt}). The widest uncertainty interval for low discharges belongs to the first period (1816-1840: dark red) for which no gaugings are available. The largest rating change arises between periods two and three, and is probably related to channel works for navigation and to the 1856 great flood consequences. After those important changes, the stage/discharge relationship stabilizes, along with a slow increase of the offsets, which may be a consequence of the filling of the channel noticed in figure \ref{fig:quantile5_both}. Offsets appear precisely estimated (i.e. posterior distributions are narrower than priors) thanks to the gaugings availability on many stage ranges. The distribution of static parameters $a$ and $c$ are presented in figure \ref{subfig:a's and c's Pt}. The $a$ parameters are precisely estimated seeing the narrow posterior distributions. However, $c$ posterior distributions are as wide as priors. This is due to the fact that $c$ priors are already very precise because they come from simplified Manning-Strickler formula for which the exponent is exactly $5/3$.

    \paragraph{}
    Eleven uncertain rating curves were computed at Beaucaire Restitution (figure \ref{subfig:RcRes}), corresponding to the stability periods derived from gaugings segmentation. They have a correct correlation with the gaugings of their respective periods. The 95\% uncertainty intervals are smaller than Pont de Beaucaire overall because of a larger number of gaugings and a smaller gaugings uncertainty. Gaugings cover a large stage range, which improves the rating curve accuracy. The first period curve is shifted with respect to the other rating curves. This is due to the scouring of the channel during the readjustment after Vallabrègues works, as described in the previous section. Offset posterior $b_1^1$ from period 1 is at least as wide as the prior in figure \ref{subfig:B'sRes}. This is because of the unavailability of low flows gaugings in the first periods. For the same reasons, the other $b_1$ offsets are poorly estimated compared to $b_2$ for which posterior distributions look particularly precise. The sustained flows of the Rhône River do not allow to explore the control of the very low water levels. Static parameters (figure \ref{subfig:bac'sRes}) appear precisely estimated, except for the 3\textsuperscript{rd} control offset $b_3$ for which the posterior distribution is as wide as the prior.

@@

        \begin{subfigure}{0.47\textwidth}
            \centering
            \includegraphics[width=\linewidth]{Figs/7e-a&csPt.pdf}
            \caption{}
            \label{subfig:a's and c's Pt}
        \end{subfigure}
        \begin{subfigure}{0.51\textwidth}
            \centering
            \includegraphics[width=\linewidth]{Figs/7f-b3_as_cs_Restit.pdf}
            \caption{}
            \label{subfig:bac'sRes}
        \end{subfigure}
        % \caption{Static parameters priors and posteriors for Pont de Beaucaire (left) and Beaucaire restitution (right)}
        \caption{Pont de Beaucaire (left) and Beaucaire Restitution (right) rating curves and relative 95\% uncertainty by respect to maxpost (first line), offsets priors and posteriors (second line) and static parameters priors and posteriors (third line). Rating curves are in logarithmic scale, solid lines represent maxpost values, grey transparent envelops represent 95\% uncertainty intervals and dots with error bars represent gaugings with 95\% uncertainty. Stable stage/discharge periods are numbered from the oldest to the latest.}
        \label{fig:RcsAndParams}
    \end{figure}

@@

    The error sources described in table \ref{tab:StageErr} are applied to both Beaucaire stations and propagated via Monte Carlo procedures. As seen in figure \ref{fig:StageErrAMAX}, measured stage is outside of stage uncertainty interval before 1840 at Pont de Beaucaire. This is due to the exponential distribution used to model measurement frequency error which is predominant here and has, by definition, only positive results. The upper uncertainty bound is sometimes 1.5 meters higher than the measured stage. Therefore, considering this source of uncertainty may have important consequences on the final results. Apart from that, the uncertainty of AMAX stages decreases over time as the measurement frequency and precision improve. The width of the 95\% uncertainty interval is 1.7 m before 1840, 0.3 m between 1840 and 1967, and 0.24 m at Beaucaire Restitution (1970-2020). The 5 m threshold above which hourly measurements were done after 1840 explains an important reduction of the uncertainty. After this date, the uncertainty is controlled by the exceedance of this 5 m threshold, the AMAX below 5 m being penalized by measurement frequency error $\delta_5$ that is considered negligible for hourly measurements made above 5m. Beaucaire Restitution uncertainty is smaller than Pont de Beaucaire uncertainty. The difference between uncertainty bounds and originally measured stages is presented in the bottom part of figure \ref{fig:StageErrAMAX}. The median of stage realisations shows some fluctuations that are mainly due to datum reference errors that are drawn every 25 years at Pont de Beaucaire.

    \begin{figure}[h!]
        \centering
        \includegraphics[width=0.8\linewidth]{Figs/8-StageErrorAMAX_BOTH.pdf}\hfill
        \caption{Top : AMAX stage uncertainty at Pont de Beaucaire (red) and Beaucaire Restitution (blue). Solid line represents measured stage. Red and blue ribbons respectively represents the 95\% uncertainty interval at Pont de Beaucaire and Beaucaire Restitution.
        Bottom : Difference between stage uncertainty bounds and originally measured stage for both stations (red and blue ribbons). Solid red lines represents the median of stage realisations and the horizontal black line at 0m stands for the measured AMAX stages. The vertical dotted line stands for the year 1840 from which hourly measurement were done above a 5m threshold. Grey band represents the years of Vallabrègues hydraulic works.}
        \label{fig:StageErrAMAX}
    \end{figure}

@@

    \begin{figure}[h!]
        \centering
        \includegraphics[width=\textwidth]{Figs/9-IC_AMAX_Both.pdf}
        \caption{AMAX maxpost discharges (black solid line) and uncertainty bounds for the three sources of streamflow uncertainties at Beaucaire (1816 - 2020). Vertical dotted lines represent rating shifts and the area of the blue dots is proportional to the number of gaugings by year.}
        \label{fig:ICtot_both}
    \end{figure}

@@

       The GEV distribution estimation procedure described in section \ref{sec:STOODS} is applied to the 205-year long AMAX discharges series at Beaucaire accounting for uncertainties. Vague priors are given to GEV parameters : flat priors for location and scale parameters, and Gaussian with zero mean and 0.2 standard deviation for the shape parameter.
       Flood quantiles results (figure \ref{fig:GEV205y}) show that streamflow uncertainty dominates for the lowest return periods, but sampling uncertainty is taking over when the return period tends toward 1000 years (see figure \ref{subfig:UKplot4cases} bottom right  for a better understanding of the respective part of each source of uncertainty in this 205 years case). The AMAX observed discharges display a large variability of streamflow uncertainties. The three largest floods of this 205 years sample (1840, 1856 and 2003, by chronological order and from the most precise to the most uncertain) illustrate this point. Thus, not considering 1840 and 1856 years could have a large effect on the estimation of the maxpost quantiles value, as well as the uncertainty bound values. This is explored next by varying the sample size.

       \begin{figure}[h!]
            \centering
            \includegraphics[width=0.7\linewidth]{Figs/10-GeV_205years.pdf}
            \caption{Flood quantiles 95\% uncertainty interval for both streamflow and total (streamflow + sampling) uncertainties. Error bars represent the AMAX observed discharges with 95\% streamflow uncertainty.}
            \label{fig:GEV205y}
        \end{figure}

        \begin{figure}[h!]
            \centering
            \begin{subfigure}{0.49\linewidth}
                \centering
                \includegraphics[width=1\linewidth]{Figs/11a-BarplotQuantiles4cases.pdf}
                \caption{}
                \label{subfig:Form4cases}
            \end{subfigure}
            \begin{subfigure}{.49\textwidth}
                \centering  
                \includegraphics[width=\linewidth]{Figs/11b-Ukplot4cases.pdf}
                \caption{}
                \label{subfig:BarplotQuantilesGev}
            \end{subfigure}

            \begin{subfigure}{0.49\textwidth}
                \centering
                \includegraphics[width=\linewidth]{Figs/11c-Shape_4cases.pdf}
                \caption{}
                \label{subfig:UKplot4cases}
            \end{subfigure}
            \begin{subfigure}{0.49\textwidth}
                \centering
                \includegraphics[width=\linewidth]{Figs/11d-Q1000SSize.pdf}
                \caption{}
                \label{subfig:SamplesQ1000}
            \end{subfigure}

        \caption{(a) Quantiles estimation for three return periods (10, 100 and 1000 years) for the four sample sizes (50, 100, 150 and 205 years). 95\% total uncertainty envelop is represented by the yellow error bars. 
        (b) Contribution of both sources of uncertainty to the total uncertainty for the four sample size cases. (c) GEV shape parameter distribution under streamflow and total uncertainty for the 4 sample sizes. (d) 1000-year flood estimations for various sample sizes from 20 to 205 years. The three grey dotted lines are representing important changes in the sample data, respectively: the change from Pont de Beaucaire to Beaucaire Restitution gauge, the inclusion of 1856 flood and the inclusion of 1840 flood to the sample.}
        \label{fig:Quantiles}
        \end{figure}

@@

        \paragraph{}
        The maxpost values of the 205 years sample in figure \ref{subfig:BarplotQuantilesGev} is higher than the 100 and 150 years samples, probably because of the inclusion of the two largest floods of the history in the 205 years sample (1840 and 1856 floods). Thus, without using those old hydrometric data although very uncertain (1816-1870), the 1000-year flood could have been 15\% lower in this specific case. The sample size impact on flood quantiles estimation is further explored in Figure \ref{subfig:SamplesQ1000}. The 1000-year flood (maxpost) and the part of both sources of uncertainties are estimated for several sample sizes, from 20 to 205 years, with a two-year step. A large reduction of sampling uncertainty (and of the total uncertainty as a consequence) appears between 20 and 100 years, along with the reduction of the maxpost value. Then, the maxpost 1000-flood value is almost constant between 100 and 160 years of sample size, until the inclusion of 1856 and 1840 floods that marks increases. The total uncertainty interval width is not much changed by those flood inclusions but the respective contribution of streamflow uncertainty appears more important.

@@

    \subsection{Strengths of the proposed approach}

>>>>

\usepackage{dsfont}
\usepackage{multirow}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{subcaption}
\usepackage{authblk}

@@

\title{Are historical stage records useful to decrease the uncertainty of flood frequency analysis ? A 200-year long case study}

\author[1]{Mathieu LUCAS}
\author[1]{Benjamin RENARD}
\author[1]{Jérôme LE COZ}
\author[1]{Michel LANG}
\author[2]{Antoine BARD}
\author[3]{Gilles PIERREFEU}

\affil[1]{INRAE Riverly, Villeurbanne}
\affil[2]{ESDB, Briançon}
\affil[3]{CNR, Lyon}

@@

    Flood frequency analysis (FFA) is a widely used method to estimate flood hazard and is affected by several sources of uncertainty. Enlarging the streamflow samples with historical continuous stage measurements can reduce sampling uncertainties, but historical flood discharges are generally more uncertain. This paper provides insights on how historical data may or may not improve design floods estimates through a chain of uncertainty estimation methods for flood frequency analysis. Uncertainties are estimated and propagated from stage and rating curves to extreme flood quantiles estimates using Monte Carlo procedures. The role of both streamflow and sampling uncertainties in design floods estimations is examined. This procedure  is applied to the 205-year long continuous discharge series of the Rhône River at Beaucaire (95 590 m\textsuperscript{2}). Total uncertainty on flood quantile is significantly reduced when the length of the series increases from 50 to 100 years. Beyond this sample size, the total uncertainty remains stable, despite 19th century flood discharges have larger uncertainties than recent flood discharges. Addition of the two largest known floods in 1840 and 1856 is increasing by 15\% the flood quantiles estimates but is almost negligible regarding the total uncertainty.

@@

    Rating curve uncertainty is also a major issue when dealing with streamflow series. Transforming stage into discharge requires calibration data (gaugings) to establish the stage/discharge relationship. Gaugings uncertainty depends on the measurement method (\citet{lecoz_quantification_2014}b and \citet{puechberty_charte_2017}). Moreover, the rating curve is also affected by uncertainties coming from the imperfection of the chosen model to represent the actual hydraulic configuration and from parameter estimation. Many methods have been proposed to quantify these uncertainties (\citet{petersen-overleir_bayesian_2009}; \citet{juston_rating_2014}; \citet{le_coz_combining_2014}; \citet{morlot_dynamic_2014}; \citet{coxon_novel_2015}; \citet{mcmillan_rating_2015}; \citet{mansanarez_rapid_2019}). A comparison of several of these methods has been recently proposed by \citet{kiang_comparison_2018}. Another important issue affecting streamflow data accuracy is rating changes. The stage/discharge relationship is frequently affected by changes coming from various origins that can be natural or anthropic, for instance: bed geometry evolution during floods or river works, aquatic vegetation growth and decay, ice cover... A regular monitoring through gaugings is essential to detect those changes \citep{ibbitt_gauging_1987}, that can be transient or sudden. Several methods have been proposed to deal with rating changes: estimating rating curves on moving temporal windows (\citet{westerberg_stage-discharge_2011} and \citet{guerrero_temporal_2012}), computing as many rating curves as there are gaugings \citep{morlot_dynamic_2014}, exploring changes in the annual minimum stages \citep{lapuszek_methods_2015}, selecting 0.5-years return period discharge as a threshold for rating changes \citep{mcmillan_impacts_2010}. More recently, \citet{darienzo_detection_2021} proposed a method based on a recursive segmentation procedure, accounting for both gaugings and rating curve uncertainties. This method has a particular interest when dealing with old and thus uncertain gaugings. Following the detection of rating shifts, rating curves should be estimated for each stability period. This task may not be straightforward, as the number of gaugings available within a stability period is not always sufficient to properly estimate the stage/discharge relationship for the whole discharge range. A common way to address this problem is to artificially repeat some gaugings from other stability periods (\citet{mcmillan_benchmarking_2012}; \citet{puechberty_charte_2017}). \citet{mansanarez_shift_2019} proposed an alternative approach to deal with this issue. They developed a stage-period-discharge (SPD) model where some parameters of the rating curve vary across periods, while the others are supposed constant. This method has the advantage of transferring information between periods to improve the rating curve estimation for all the stable stage/discharge periods, even when few gaugings are available.

@@

    \paragraph{}
    Once the stability periods have been identified, the next step is to estimate the rating curve associated with each period. \citet{mansanarez_shift_2019} developed a stage-period-discharge (SPD) model "based on the physical interpretation of changes in the stage-discharge relation across a series of stability periods". The SPD model is based on the BaRatin model \citep{le_coz_combining_2014}.
    \paragraph{} BaRatin uses the Bayesian paradigm to estimate the parameters of the general rating curve equation, a combination of power equations: $Q = a(h-b)^c$, where $Q$ is the discharge in m\textsuperscript{3}/s, the flow depth is the difference between the stage $h$ (in m) and an offset $b$, and $a$ and $c$ being the coefficient and the exponent of the power function. The rating curve equation is deduced from a hydraulic analysis of the gauging station, aimed at identifying the main hydraulic controls governing the stage-discharge relation. The multiple controls can be activated successively or simultaneously. Bayesian inference allows deriving the posterior distribution of rating curve parameters by combining hydraulic information (priors for parameters of each hydraulic controls) and information from gaugings with uncertainty (likelihood).
    \paragraph{}
    The SPD model is estimating the rating curves of each stability based on the same principle, and is considering that several identified parameters vary in time, throughout the stability periods. An important step is the identification of those varying parameters based on hydraulic analysis of the site. Generally, offsets or channel widths are suspected to change. The so-called local changes are supposed to affect the lowest control only (for instance the movement of the controlling riffle), whereas the global changes affect various controls at the same time (for instance, the scouring or filling of the channel, affecting low flows and main channel offsets at the same time). One way to follow the evolution of riverbed elevation and to help specifying priors for rating curve vertical evolution is to study the long-term evolution of the yearly lowest stages, as described by \citet{lapuszek_methods_2015}. See \citet{mansanarez_shift_2019} for a detailed description of prior specification for time-varying rating curves.

@@

    \paragraph{}    
    Thirteen shift times are detected at Beaucaire Restitution. As can be seen in figure \ref{fig:SegmBoth} (right), gauging frequency is far higher than for Pont de Beaucaire station (except during the 5 first years), resulting in a better determination of the rating shift times. 
    Due to the lack of gaugings at the beginning of the series, only one rating shift is detected but many shifts potentially took place in those first four years as morphological readjustments and dredging occurred (see previous section). This first shift is assumed to be assigned to the first important flood of the station in 1976, after which the channel is stabilized. The following shift occurred during 1994 flood, one of the biggest of the station. The most notable flood at Beaucaire Restitution occurred in 2003 (11 500 m\textsuperscript{3}/s, with a return period of about 100 years according to \citet{medd_debit_2005}). Unsurprisingly, the stage/discharge relationship is considerably disrupted by this event, as reflected by the six rating shifts detected from 2003 to 2005. Two out of these six shifts were discarded because the shift amplitude is considered minor based on further analysis of the corresponding rating curve change. The largest flood within posterior pdf of those shifts is almost always corresponding to the 2003 flood. This is also the case for 2005, 2008 and 2009 shifts, for which the posterior pdf spans many years and include 2003 flood. Therefore, the shift dates are assumed to be located to the maxpost shift times, as several shifts can not be located at the same date. Lastly, the last shift of 2019 is also discarded because the shift amplitude is considered minor based on further analysis. Finally, ten rating shifts were retained. This brings the number of stability periods to eleven for Beaucaire Restitution, as seen in table \ref{tab:ShiftDates}.

@@

    \paragraph{}
    Uncertain rating curves are estimated using \citet{mansanarez_shift_2019} SPD model, for each stability period detected previously. For Pont de Beaucaire, this leads to ten rating curves that show a good adequacy with gaugings (figure \ref{subfig:RcPt}). The evolution of the low flows control offset ($b_1$) gives indications on the evolution of bed elevation (figure \ref{subfig:B'sPt}). Important changes occurred during the third and fourth stability periods with a successive increase and decrease of the offset. Those changes may be related to the channel works that occurred during the end of the XIX\textsuperscript{th} Century. Afterwards, the offset is stabilized, along with a slight increase trend which may be a consequence of the filling of the channel noticed in figure \ref{fig:quantile5_both}. The widest uncertainty interval belongs to the first period (1816-1840: dark red) for which no gaugings are available (figure \ref{subfig:RcPt}). The expected range of rating curve uncertainties for floods discharges (above 6 m) varies from around 20\% for the first period, to less than 10\% after 1840. Static parameters are precisely estimated and are presented in figure figure \ref{subfig:a's and c's Pt}. The $c$ posterior distributions are as wide as priors, due to the fact that $c$ priors are already very precise because they come from simplified Manning-Strickler formula for which the exponent is exactly $5/3$.
    % Static parameters $a$ and $c$ (figure \ref{subfig:a's and c's Pt}). The $a$ parameters are precisely estimated seeing the narrow posterior distributions. However, $c$ posterior distributions are as wide as priors. This is due to the fact that $c$ priors are already very precise because they come from simplified Manning-Strickler formula for which the exponent is exactly $5/3$.

    \paragraph{}
    Eleven uncertain rating curves were computed at Beaucaire Restitution (figure \ref{subfig:RcRes}). The rating curve uncertainty intervals are smaller than Pont de Beaucaire overall because of a larger number of gaugings and a smaller gaugings uncertainty: around 5\% of uncertainty is expected for floods above 6 m. However, low stages uncertainty is greater than at Pont de Beaucaire, because of the unavailability of low flows gaugings. The sustained flows of the Rhône River limits the exploration of the low flows hydraulic control.
    The first period rating curve (dark red) is shifted with respect to the other curves due to the channel readjustments and dredging operations after Vallabrègues works (1967-1970). The first control offsets ($b_1$) are globally decreasing over time, showing a slight scouring trend of the channel (figure \ref{subfig:B'sRes}). Static parameters (figure \ref{subfig:bac'sRes}) appear precisely estimated, except for the 3\textsuperscript{rd} control offset $b_3$ for which the posterior distribution is as wide as the prior.

@@

        \begin{subfigure}{0.39\textwidth}
            \centering
            \includegraphics[width=\linewidth]{Figs/7e-a&csPt.pdf}
            \caption{}
            \label{subfig:a's and c's Pt}
        \end{subfigure}
        \begin{subfigure}{0.41\textwidth}
            \centering
            \includegraphics[width=\linewidth]{Figs/7f-b3_as_cs_Restit.pdf}
            \caption{}
            \label{subfig:bac'sRes}
        \end{subfigure}
        % \caption{Static parameters priors and posteriors for Pont de Beaucaire (left) and Beaucaire restitution (right)}
        \caption{Pont de Beaucaire (a) and Beaucaire Restitution (b) rating curves and relative 95\% uncertainty by respect to maxpost, offsets priors and posteriors (c and d) and static parameters priors and posteriors (e and f). Rating curves are in logarithmic scale, solid lines represent maxpost values, grey transparent envelops represent 95\% uncertainty intervals and dots with error bars represent gaugings with 95\% uncertainty. Stable stage/discharge periods are numbered from the oldest to the latest.}
        \label{fig:RcsAndParams}
    \end{figure}

@@

    The error sources described in table \ref{tab:StageErr} are applied to both Beaucaire stations and propagated via Monte Carlo procedures. As seen in figure \ref{fig:StageErrAMAX}, measured stage is outside of stage uncertainty interval before 1840 at Pont de Beaucaire. This is due to the exponential distribution used to model measurement frequency error which is predominant here and has, by definition, only positive results. The upper uncertainty bound is sometimes 1.5 meters higher than the measured stage. Therefore, considering this source of uncertainty may have important consequences on the final results. Apart from that, the uncertainty of AMAX stages decreases over time as the measurement frequency and precision improve. The width of the 95\% uncertainty interval is 1.7 m before 1840, 0.3 m between 1840 and 1967, and 0.24 m at Beaucaire Restitution (1970-2020). The 5 m threshold above which hourly measurements were done after 1840 explains an important reduction of the uncertainty. After this date, the uncertainty is controlled by the exceedance of this 5 m threshold, the AMAX below 5 m being penalized by measurement frequency error $\delta_5$ that is considered negligible for hourly measurements made above 5 m. Beaucaire Restitution uncertainty is smaller than Pont de Beaucaire uncertainty. The difference between uncertainty bounds and originally measured stages is presented in the bottom part of figure \ref{fig:StageErrAMAX}. The median of stage realisations shows some fluctuations that are mainly due to datum reference errors that are drawn every 25 years at Pont de Beaucaire.

    \begin{figure}[h!]
        \centering
        \includegraphics[width=0.9\linewidth]{Figs/8-StageErrorAMAX_BOTH.pdf}\hfill
        \caption{Top : AMAX stages and uncertainty at Beaucaire (1816-2020). Bottom : Difference between stage uncertainty bounds and originally measured stage.}
        \label{fig:StageErrAMAX}
    \end{figure}

@@

    \begin{figure}[h!]
        \centering
        \includegraphics[width=\textwidth]{Figs/9-IcAndAMAX.pdf}
        \caption{Relative uncertainty of AMAX discharges by respect to the maxpost discharges (black solid line) for the three sources of streamflow uncertainties, at Beaucaire (1816 - 2020). Vertical dotted lines represent rating shifts and the area of the blue dots is proportional to the number of gaugings by year.}
        \label{fig:ICtot_both}
    \end{figure}

@@

       The GEV distribution estimation procedure described in section \ref{sec:STOODS} is applied to the 205-year long AMAX discharges series at Beaucaire accounting for uncertainties. Vague priors are given to GEV parameters : flat priors for location and scale parameters, and Gaussian with zero mean and 0.2 standard deviation for the shape parameter.
       Flood quantiles results (figure \ref{subfig:GEV205y}) show that streamflow uncertainty dominates for the lowest return periods, but sampling uncertainty is taking over when the return period tends toward 1000 years (see figure \ref{subfig:UKplot4cases} bottom right  for a better understanding of the respective part of each source of uncertainty in this 205 years case). The AMAX observed discharges display a large variability of streamflow uncertainties. The three largest floods of this 205 years sample (1840, 1856 and 2003, by chronological order and from the most precise to the most uncertain) illustrate this point. Thus, not considering 1840 and 1856 years could have a large effect on the estimation of the maxpost quantiles value, as well as the uncertainty bound values. This is explored next by varying the sample size.

       % \begin{figure}[h!]
       %      \centering
       %      \includegraphics[width=0.7\linewidth]{Figs/10-GeV_205years.pdf}
       %      \caption{Flood quantiles 95\% uncertainty interval for both streamflow and total (streamflow + sampling) uncertainties. Error bars represent the AMAX observed discharges with 95\% streamflow uncertainty.}
       %      \label{fig:GEV205y}
       %  \end{figure}

        \begin{figure}[p]
            \centering
            \begin{subfigure}{0.99\textwidth}
                \centering
                \includegraphics[width=1\linewidth]{Figs/10-GeV_205years.pdf}
                \caption{}
                \label{subfig:GEV205y}
            \end{subfigure}

            \centering
            \begin{subfigure}{0.49\linewidth}
                \centering
                \includegraphics[width=1\linewidth]{Figs/11a-BarplotQuantiles4cases.pdf}
                \caption{}
                \label{subfig:BarplotQuantilesGev}
            \end{subfigure}
            \begin{subfigure}{.49\textwidth}
                \centering  
                \includegraphics[width=\linewidth]{Figs/11b-Ukplot4cases.pdf}
                \caption{}
                \label{subfig:UKplot4cases}
            \end{subfigure}

            \begin{subfigure}{0.47\textwidth}
                \centering
                \includegraphics[width=\linewidth]{Figs/11c-Shape_4cases.pdf}
                \caption{}
                \label{subfig:Form4cases}
            \end{subfigure}
            \begin{subfigure}{0.52\textwidth}
                \centering
                \includegraphics[width=\linewidth]{Figs/11d-Q1000SSize.pdf}
                \caption{}
                \label{subfig:SamplesQ1000}
            \end{subfigure}

        \caption{(a) Flood quantiles at Beaucaire. Error bars represent the AMAX observed discharges with 95\% streamflow uncertainty.
        (b) Quantiles estimation of three return periods for four sample sizes. 95\% total uncertainty envelop is represented by the yellow error bars. 
        (c) Contribution of both sources of uncertainty to the total uncertainty for four sample size. 
        (d) GEV shape parameter distribution considering streamflow or total uncertainty for four sample sizes. (e) 1000-year flood estimations for various sample sizes. Grey dotted lines are representing important changes in the sample data (the change from Pont de Beaucaire to Beaucaire Restitution gauge en 1967, the inclusion of 1856 flood and the inclusion of 1840 flood to the sample.}
        \label{fig:Quantiles}
        \end{figure}

@@

        \paragraph{}
        The maxpost values of the 205 years sample is higher than the 100 and 150 years samples (figure \ref{subfig:BarplotQuantilesGev}), probably because of the inclusion of the two largest floods of the history in the 205 years sample (1840 and 1856 floods). Thus, without using those old hydrometric data although very uncertain (1816-1870), the 1000-year flood could have been 15\% lower in this specific case. The sample size impact on flood quantiles estimation is further explored in Figure \ref{subfig:SamplesQ1000}. The 1000-year flood (maxpost) and the part of both sources of uncertainties are estimated for several sample sizes, from 20 to 205 years, with a two-year step. A large reduction of sampling uncertainty (and of the total uncertainty as a consequence) appears between 20 and 100 years, along with the reduction of the maxpost value. Then, the maxpost and uncertainty bounds are constant between 100 and 160 years of sample size, until the increase due to the inclusion of 1856 and 1840 major floods. The total uncertainty interval width is not much changed by those flood inclusions but the respective contribution of streamflow uncertainty appears more important.

@@

    \subsection{Consequences of including several uncertainty sources in FFA with historical stage records}
